---
title: "FinPred"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{FinPred}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction


## Installation

```{r setup}
library(FinPred)
```


## Using the Package


## References
Box, G. E. P., Jenkins, G. M., & Reinsel, G. C. (1994). Time Series Analysis: Forecasting and Control. Prentice Hall. 

Breiman, L. (2001). Machine Learning, 45(1), 5–32. https://doi.org/10.1023/a:1010933404324 

Chen, T., & Guestrin, C. (2016). XGBoost. Proceedings of the 22nd ACM SIGKDD International Conference on Knowledge Discovery and Data Mining. https://doi.org/10.1145/2939672.2939785 

Chen T, He T, Benesty M, Khotilovich V, Tang Y, Cho H, Chen K, Mitchell R, Cano I,
  Zhou T, Li M, Xie J, Lin M, Geng Y, Li Y, Yuan J (2022). _xgboost: Extreme
  Gradient Boosting_. R package version 1.6.0.1,
  <https://CRAN.R-project.org/package=xgboost>.

Cortes, C., & Vapnik, V. (1995). Support-Vector Networks. Machine Learning, 20(3), 273–297. https://doi.org/10.1007/bf00994018 

Csardi G, Nepusz T: The igraph software package for complex network research,
  InterJournal, Complex Systems 1695. 2006. https://igraph.org

Komsta L, Novomestky F (2022). _moments: Moments, Cumulants, Skewness, Kurtosis
  and Related Tests_. R package version 0.14.1,
  <https://CRAN.R-project.org/package=moments>.

Liaw A., Wiener  M. (2002). Classification and Regression by randomForest. R
  News 2(3), 18--22.

Meyer D, Dimitriadou E, Hornik K, Weingessel A, Leisch F (2022). _e1071: Misc
  Functions of the Department of Statistics, Probability Theory Group (Formerly:
  E1071), TU Wien_. R package version 1.7-11,
  <https://CRAN.R-project.org/package=e1071>.

R Core Team (2022). R: A language and environment for statistical computing. R
  Foundation for Statistical Computing, Vienna, Austria. URL
  https://www.R-project.org/.

Ryan JA, Ulrich JM (2022). _quantmod: Quantitative Financial Modelling Framework_.
  R package version 0.4.20, <https://CRAN.R-project.org/package=quantmod>.

Tsay, R. S. (2010). Analysis of Financial Time Series. John Wiley & Sons.


